---
title: "birth order"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
twins2013_matched <- read.csv("~/Desktop/PhD/2021 Spring/STAT 566/project/data/twins_2013_matched.csv")
twins2013_matched <- twins2013_matched%>%select(-X, -X1)

missing.wt<-twins2013_matched$number[matched_all$dbwt==9999]
twins2013_matched<-twins2013_matched%>%
  filter(!(number%in%missing.wt),cig_1!=99)%>%
  mutate(preg.condi=str_detect(paste0(rf_diab,rf_gest,rf_phyp,rf_ghyp,rf_eclam,rf_ppterm,rf_ppoutc,rf_cesar),"Y"))
```

# Determine Birth Order
```{r}
#create subsets for twins, triplets, quadruplets, quintuplets
table(twins2013_matched$n_children)
twins_2013_twins<-twins2013_matched%>%filter(n_children==2)
twins_2013_trip<-twins2013_matched%>%filter(n_children==3)
twins_2013_quad<-twins2013_matched%>%filter(n_children==4)
twins_2013_quin<-twins2013_matched%>%filter(n_children==5)
# Note: a part of them are not orderable because of how they coded the variable that 8 indicates 8 or more kids

#look at quintuplets 
twins_2013_quin%>%select(number, lbo, tbo, death) #tbo doesn't make perfect sense to me
twins_2013_quin%>%group_by(number)%>%summarise(orderable1=length(unique(lbo))==5,
                                               orderable2=(max(lbo)-min(lbo)==4),
                                               orderable=orderable1&orderable2) #the only problematic one is 22197
twins_2013_quin_ord<-twins_2013_quin%>%filter(number!=22197)

#look at quadrauplets
twins_2013_quad%>%select(number, lbo, tbo, death) #tbo doesn't make perfect sense to me
orderable_quad<-twins_2013_quad%>%group_by(number)%>%summarise(orderable1=length(unique(lbo))==4,
                                               orderable2=(max(lbo)-min(lbo)==3),
                                               orderable=orderable1&orderable2)
table(orderable_quad$orderable)
orderable_num<-orderable_quad$number[orderable_quad$orderable]
twins_2013_quad%>%filter(!(number%in%orderable_num))%>%select(number, lbo, tbo, death)
twins_2013_quad_ord<-twins_2013_quad%>%filter(number%in%orderable_num)
length(unique(twins_2013_quad_ord$number))==length(orderable_num)

#look at triplets
twins_2013_trip%>%select(number, lbo, tbo, death)
orderable_trip<-twins_2013_trip%>%group_by(number)%>%summarise(orderable1=length(unique(lbo))==3,
                                               orderable2=(max(lbo)-min(lbo)==2),
                                               orderable=orderable1&orderable2)
table(orderable_trip$orderable)
orderable_num<-orderable_trip$number[orderable_trip$orderable]
twins_2013_trip%>%filter(!(number%in%orderable_num))%>%select(number, lbo, tbo, death)
twins_2013_trip_ord<-twins_2013_trip%>%filter(number%in%orderable_num)
length(unique(twins_2013_trip_ord$number))==length(orderable_num)


#look at twins
twins_2013_twins%>%select(number, lbo, tbo, death)
orderable_twin<-twins_2013_twins%>%group_by(number)%>%summarise(orderable1=length(unique(lbo))==2,
                                               orderable2=(max(lbo)-min(lbo)==1),
                                               orderable=orderable1&orderable2)
table(orderable_twin$orderable)
orderable_num<-orderable_twin$number[orderable_twin$orderable]
twins_2013_twins%>%filter(!(number%in%orderable_num))%>%select(number, lbo, tbo, death)
twins_2013_twin_ord<-twins_2013_twins%>%filter(number%in%orderable_num)
length(unique(twins_2013_twin_ord$number))==length(orderable_num)

#combine all orderable data
twins_2013_ord<-rbind(twins_2013_quin_ord, twins_2013_quad_ord, twins_2013_trip_ord, twins_2013_twin_ord)
nrow(twins_2013_quad_ord)+nrow(twins_2013_quin_ord)+nrow(twins_2013_trip_ord)+nrow(twins_2013_twin_ord)==nrow(twins_2013_ord)

#percent of orderable data out of matched data
nrow(twins_2013_ord)/nrow(twins2013_matched)
length(unique(twins_2013_ord$number))/length(unique(twins2013_matched$number))

#add orderable variable in twins2013_matched
allord_num<-unique(twins_2013_ord$number)
twins2013_matched<-twins2013_matched%>%mutate(orderable=ifelse(number%in%allord_num,1,0))
table(twins2013_matched$orderable)
```

# if missing birth order info is associated with some covariates

```{r}
death_count<-twins2013_matched%>%group_by(number)%>%summarize(n_death=sum(death),n_death1=ifelse(n_death>0,1,0))
twins2013_matched<-twins2013_matched%>%left_join(death_count, by="number")

#pick one out of each pair
library(splitstackshape)
stratified_sample<-stratified(twins2013_matched, "number", 1)
nrow(stratified_sample)
length(unique(twins2013_matched$number))

#filter out lbo=8
stratified_no8<-stratified_sample%>%filter(lbo!=8)
```

Some ideas:
  - use the stratified sample to see if some general characteristic of parents is associated with missing
  - if having a death is associated with missing
  - how to deal with lbo=8? omit? done
  - further check if lbo doesn't match tbo
  - consistent missingness - coded 9 or 99 or U

```{r}
death_missing<-glm(orderable ~ n_death, family="binomial", data=stratified_no8)
summary(death_missing)
exp(coef(death_missing))

death_missing1<-glm(orderable ~ n_death1, family="binomial", data=stratified_no8)
summary(death_missing1)
exp(coef(death_missing1))
1/0.7362388
```

```{r}
mom_missing<-glm(orderable ~ meduc+mracerec+fracerec+mager41, family="binomial", data=stratified_no8)
summary(mom_missing)

table(as.factor(stratified_no8$meduc))
educ_missing<-glm(orderable ~ as.factor(meduc), family="binomial", data=stratified_no8)
summary(educ_missing)

table(as.factor(stratified_no8$mracerec))
mrace_missing<-glm(orderable ~ as.factor(mracerec), family="binomial", data=stratified_no8)
summary(mrace_missing)
regress("odds", orderable ~ as.factor(mracerec), data=stratified_no8)

table(as.factor(stratified_no8$fracerec))
frace_missing<-glm(orderable ~ as.factor(fracerec), family="binomial", data=stratified_no8)
summary(frace_missing)

mage_missing<-glm(orderable ~ mager41, family="binomial", data=stratified_no8)
summary(mage_missing)
```

```{r}
ggplot(data=stratified_no8)+
  geom_bar(aes(x=as.factor(meduc),fill=as.factor(orderable)), position = "fill")

ggplot(data=stratified_no8)+
  geom_bar(aes(x=as.factor(mracerec),fill=as.factor(orderable)), position = "fill")+
  scale_x_discrete(labels = c("White", "Black", "American Indian / Alaskan Native", "Asian / Pacific Islander"))+
  labs(x="mother's race", title = "Mother's racing and birth order info")+
  scale_fill_discrete(name="Having birth order info",
                         labels=c("No", "Yes"))

```


```{r}
temp<-twins2013_matched%>%filter(lbo==8)
nrow(temp)
length(unique(temp$number))
nrow(temp)/(nrow(twins2013_matched)-nrow(twins_2013_ord))

temp<-twins2013_matched%>%filter(lbo==9)
nrow(temp)
length(unique(temp$number))
```

## Analysis: Does birth order have an impact on birth weight?

level 1: $bw=b0+b1*birth_order+b2*apgar5+b3*sex$

level 2: $b0=a0+a1*mager41+a2*mracerec+a3*fracerec+a4*cig_1+a5*wtgain+a6*preg.condi$

```{r}
order_assign<-twins_2013_ord%>%group_by(number)%>%summarise(lbo=lbo,order=lbo-min(lbo)+1)
twins_2013_ord<-twins_2013_ord%>%left_join(order_assign, by=c("lbo", "number"))

set.seed(4)
random.num<-sample(unique(matched_all$number), size=10000, replace = F)
random.twins_2013_ord<-twins_2013_ord%>%filter(number%in% random.num)

mixed.lmer1 <- lmer(dbwt ~ order + mager41+mracerec+fracerec+cig_1+wtgain+sex+apgar5+preg.condi +(1|number), data = random.twins_2013_ord)
summary(mixed.lmer1)

mixed.lmer2 <- lmer(dbwt ~ order + mager41+mracerec+fracerec+cig_1+wtgain+sex+apgar5+preg.condi +(order|number), data = twins_2013_ord)
summary(mixed.lmer2)

mixed.lmer3 <- lmer(dbwt ~ order + mager41+mracerec+fracerec+cig_1+wtgain+sex+apgar5+preg.condi +(1|number), data = twins_2013_ord)
summary(mixed.lmer3)

mixed.lmer.uncon <- lmer(dbwt ~ 1 +(1|number), data = random.twins_2013_ord)
summary(mixed.lmer.uncon)

mixed.lmer.uncon2 <- lmer(dbwt ~ 1 +(1|number), data = twins_2013_ord)
summary(mixed.lmer.uncon2)

anova(mixed.lmer.uncon, mixed.lmer1)
```



























