---
title: "Birth weight"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(splitstackshape)
matched0<-read.csv("~/Desktop/PhD/2021 Spring/STAT 566/project/twins-analysis/data/twins2013_matched_children.csv")

matched_twin<-matched0%>%
  filter(n_children==2)%>%
  select(-X.1, -X)

missing.wt<-matched_twin$number[matched_twin$dbwt==9999]

matched_twin<-matched_twin%>%
  filter(!(number%in%missing.wt))

matched_twin%>%filter(dbwt==9999)

matched_all<-matched0%>%
  select(-X.1, -X)

missing.wt<-matched_all$number[matched_all$dbwt==9999]

matched_all<-matched_all%>%
  filter(!(number%in%missing.wt))
matched_all%>%filter(dbwt==9999)

matched_diff<-matched_diff%>%
  filter(cig_1!=99)%>%
  mutate(preg.condi=str_detect(paste0(rf_diab,rf_gest,rf_phyp,rf_ghyp,rf_eclam,rf_ppterm,rf_ppoutc,rf_cesar),"Y"))

matched_all<-matched_all%>%
  filter(cig_1!=99)%>%
  mutate(preg.condi=str_detect(paste0(rf_diab,rf_gest,rf_phyp,rf_ghyp,rf_eclam,rf_ppterm,rf_ppoutc,rf_cesar),"Y"))
```

# Question 1: Does mother’s health have an impact on the difference in birth weight of twin pairs?

## Part 1: Birth weight difference for twins

```{r}
wtdiff<-matched_twin%>%
  group_by(number)%>%
  summarise(wt_diff=max(dbwt)-min(dbwt), 
            sex_pair=paste0(sex[1],sex[2]), 
            sex_pairb=ifelse(sex_pair=="FF", 0, 
                             ifelse(sex_pair %in% c("MF","FM"),1,2)),)
matched_diff<-matched_twin%>%
  left_join(wtdiff, by="number")

sort(matched_diff$wt_diff)
```

```{r}
#Some EDA
ggplot(data=matched_diff)+
  geom_point(aes(x=mager41, y=wt_diff))

head(sort(matched_diff$wt_diff, decreasing = TRUE))

matched_diff%>%filter(wt_diff==5820)


ggplot(data=matched_diff)+
  geom_boxplot(aes(x=as.factor(mracerec), y=wt_diff))+
  ylim(0,1000)
  

ggplot(data=matched_diff)+
  geom_boxplot(aes(x=as.factor(fracerec), y=wt_diff))+
  ylim(0,1000)


ggplot(data=matched_diff)+
  geom_boxplot(aes(x=as.factor(cig_1), y=wt_diff))+
  ylim(0,1000)


ggplot(data=matched_diff)+
  geom_boxplot(aes(x=as.factor(wtgain), y=wt_diff))+
  ylim(0,1000)

ggplot(data=matched_diff)+
  geom_boxplot(aes(x=sex_pair, y=wt_diff))+
  ylim(0,1000)

ggplot(data=matched_diff)+
  geom_boxplot(aes(x=as.factor(sex_pairb), y=wt_diff))+
  ylim(0,1000)

ggplot(data=matched_diff)+
  geom_boxplot(aes(x=as.factor(mar), y=wt_diff))+
  ylim(0,1000)

ggplot(data=matched_diff)+
  geom_boxplot(aes(x=as.factor(restatus), y=wt_diff))+
  ylim(0,1000)

ggplot(data=matched_diff)+
  geom_boxplot(aes(x=preg.condi, y=wt_diff))+
  ylim(0,1000)

```


```{r}
# orderable_twin<-matched_diff%>%
#   group_by(number)%>%
#   summarise(orderable1=length(unique(lbo))==2,
#          orderable2=(max(lbo)-min(lbo)==1),
#          orderable=orderable1&orderable2)
# 
# table(orderable_twin$orderable)
# 
# orderable_num<-orderable_twin$number[orderable_twin$orderable]
# 
# matched_diff_ord<-matched_diff%>%
#   mutate(orderable=ifelse(number%in%orderable_num,1,0)) #actually don't need all steps before since not interested in order

matched_diff_ord_stratified<-stratified(matched_diff, "number", 1)

hist(matched_diff_ord_stratified$wt_diff, xlim = c(0,500), breaks = 3000)

lm1<-lm(wt_diff~mager41+mracerec+fracerec+cig_1+wtgain+sex_pair+mar+preg.condi, data = matched_diff_ord_stratified)
summary(lm1)

lm2<-lm(wt_diff~mager41+mracerec+wtgain+sex_pair+preg.condi, data = matched_diff_ord_stratified)
summary(lm2)

# matched_ord_1<-matched_ord%>%
#   group_by(number)%>%
#   summarise(dbwt=dbwt,order=ifelse(orderable==1, lbo-min(lbo)+1, row_number()))
# table((matched_ord_1%>%group_by(number)%>%summarise(check=(max(order)==2)))$check)
# 
# matched_with_order <- matched_ord%>%
#   left_join(matched_ord_1, by=c("number", "dbwt"))
#   
# try_wide<-matched_with_order[1:100, ]%>%
#   spread(key=order, value=apgar5)




```


### Repeat the analysis for g-g, g-b, b-b


```{r}
matched_diff_ord_stratified_FF<-matched_diff_ord_stratified%>%filter(sex_pairb==0)
matched_diff_ord_stratified_MF<-matched_diff_ord_stratified%>%filter(sex_pairb==1)
matched_diff_ord_stratified_MM<-matched_diff_ord_stratified%>%filter(sex_pairb==2)
matched_diff_ord_stratified_homo<-matched_diff_ord_stratified%>%filter(sex_pairb!=1)
```


```{r}
lm1_FF<-lm(wt_diff~mager41+mracerec+fracerec+cig_1+wtgain+mar+preg.condi, data = matched_diff_ord_stratified_FF)
summary(lm1_FF)

lm2_FF<-lm(wt_diff~mager41+mracerec+wtgain+preg.condi, data = matched_diff_ord_stratified_FF)
summary(lm2_FF)

lm1_MF<-lm(wt_diff~mager41+mracerec+fracerec+cig_1+wtgain+mar+preg.condi, data = matched_diff_ord_stratified_MF)
summary(lm1_MF)

lm2_MF<-lm(wt_diff~mager41+mracerec+wtgain+preg.condi, data = matched_diff_ord_stratified_MF)
summary(lm2_MF)


lm1_MM<-lm(wt_diff~mager41+mracerec+fracerec+cig_1+wtgain+mar+preg.condi, data = matched_diff_ord_stratified_MM)
summary(lm1_MM)

lm2_MM<-lm(wt_diff~mager41+mracerec+wtgain+preg.condi, data = matched_diff_ord_stratified_MM)
summary(lm2_MM)
#mother's race is not less significant

lm1_homo<-lm(wt_diff~mager41+mracerec+fracerec+cig_1+wtgain+mar+preg.condi, data = matched_diff_ord_stratified_homo)
summary(lm1_homo)

lm2_homo<-lm(wt_diff~mager41+mracerec+wtgain+preg.condi, data = matched_diff_ord_stratified_homo)
summary(lm2_homo)


```



## Part 2: Mean birth weight for all non-singletons

```{r}
table(matched_all$n_children)
wtmean<-matched_all%>%
  group_by(number)%>%
  summarise(wt_mean=mean(dbwt), 
            sex_pair=str_replace_all(paste0(sex[1],sex[2],sex[3],sex[4],sex[5]),"NA","" ),
            sex_pairb=ifelse(sex_pair%in%c("FF","FFF","FFFF","FFFFF"), 0, 
                             ifelse(sex_pair %in% c("MM","MMM","MMMM","MMMMM"),2,1))
            )

matched_mean<-matched_all%>%
  left_join(wtmean, by="number")
```

```{r}
#Some EDA
ggplot(data=matched_mean)+
  geom_point(aes(x=mager41, y=wt_mean))

head(sort(matched_diff$wt_diff, decreasing = TRUE))

matched_diff%>%filter(wt_diff==5820)


ggplot(data=matched_mean)+
  geom_boxplot(aes(x=as.factor(mracerec), y=wt_mean))
  

ggplot(data=matched_mean)+
  geom_boxplot(aes(x=as.factor(fracerec), y=wt_mean))


ggplot(data=matched_mean)+
  geom_boxplot(aes(x=as.factor(cig_1), y=wt_mean))


ggplot(data=matched_mean)+
  geom_boxplot(aes(x=as.factor(wtgain), y=wt_mean))


ggplot(data=matched_mean)+
  geom_boxplot(aes(x=as.factor(sex_pairb), y=wt_mean))

ggplot(data=matched_mean)+
  geom_boxplot(aes(x=as.factor(mar), y=wt_mean))

ggplot(data=matched_mean)+
  geom_boxplot(aes(x=as.factor(restatus), y=wt_mean))

ggplot(data=matched_mean)+
  geom_boxplot(aes(x=preg.condi, y=wt_mean))

```


```{r}
matched_mean_stratified<-stratified(matched_mean, "number", 1)

hist(matched_mean_stratified$wt_mean,  breaks = 300)

lm1<-lm(wt_mean~mager41+mracerec+fracerec+cig_1+wtgain+as.factor(sex_pairb)+mar+preg.condi, data = matched_mean_stratified)
summary(lm1)

```



### Repeat the analysis for all g, all b, mix

```{r}
matched_mean_stratified_allF<-matched_mean_stratified%>%filter(sex_pairb==0)
matched_mean_stratified_FandM<-matched_mean_stratified%>%filter(sex_pairb==1)
matched_mean_stratified_allM<-matched_mean_stratified%>%filter(sex_pairb==2)
matched_mean_stratified_homo<-matched_mean_stratified%>%filter(sex_pairb!=1)
```


```{r}
lm1_allF<-lm(wt_mean~mager41+mracerec+fracerec+cig_1+wtgain+mar+preg.condi, data = matched_mean_stratified_allF)
summary(lm1_allF)


lm1_FandM<-lm(wt_mean~mager41+mracerec+fracerec+cig_1+wtgain+mar+preg.condi, data = matched_mean_stratified_FandM)
summary(lm1_FandM)


lm1_allM<-lm(wt_mean~mager41+mracerec+fracerec+cig_1+wtgain+mar+preg.condi, data = matched_mean_stratified_allM)
summary(lm1_allM) #cig_1 is not significant


lm1_homo<-lm(wt_mean~mager41+mracerec+fracerec+cig_1+wtgain+mar+preg.condi, data = matched_mean_stratified_homo)
summary(lm1_homo)

```


# Question 2: Does birth weight have an impact on mortality rate?


level 1: $y=b0+b1*bw+b2*apgar5+b3*sex$

level 2: $b0=a0+a1*mager41+a2*mracerec+a3*fracerec+a4*cig_1+a5*wtgain$
         
```{r}
set.seed(4)
random.num<-sample(unique(matched_all$number), size=5000, replace = F)
random.matched_all<-matched_all%>%filter(number%in% random.num)
library(lme4)
mixed.lmer1 <- glmer(death ~ dbwt + mager41+mracerec+fracerec+cig_1+wtgain+sex+apgar5 +(1|number), data = random.matched_all, family = "binomial")
summary(mixed.lmer1)

random.matched_all<-random.matched_all%>%mutate(low.bw=ifelse(dbwt<2500,1,0), smoking=ifelse(cig_1>0,1,0))

mixed.lmer2 <- glmer(death ~ low.bw +smoking+ low.bw*smoking + mager41+mracerec+sex+(1|number), data = random.matched_all, family = "binomial")
summary(mixed.lmer2)

mixed.lmer3 <- glmer(death ~ as.factor(low.bw) +as.factor(smoking)+ as.factor(low.bw) *as.factor(smoking) + mager41+mracerec+fracerec+wtgain+sex+apgar5 + preg.condi +(1|number), data = random.matched_all, family = "binomial")
summary(mixed.lmer3)

mixed.lmer4 <- glmer(death ~ as.factor(low.bw) +as.factor(smoking)+ as.factor(low.bw) *as.factor(smoking) + mager41+mracerec+sex+apgar5 +(1|number), data = random.matched_all, family = "binomial")
summary(mixed.lmer4)

mean(random.matched_all$low.bw)

matched_all<-matched_all%>%mutate(low.bw=ifelse(dbwt<2500,1,0), smoking=ifelse(cig_1>0,1,0))
mean(matched_all$low.bw)
```

mean birth weight and mortality rate within pair?


# PC algorithm

