---
title: "EDA"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

Exploratory data analysis

Install and load dependencies
```{r}
list.of.packages <- c("knitr", "tidyverse", "ggplot2", "Rcpp", "gridExtra")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
library(gridExtra)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r}
library(tidyverse)
# Load data
matched_children_df<-read.csv("./data/twins2013_matched_children.csv")
```

Clean data to only look at twins
```{r}
# Separate out twins
print('Original  dataset had rows:' + nrow(matched_children_df))
twins_only <- filter(matched_children_df, n_children == 2)
print('Only twins:' + nrow(twins_only))
```

Create dataset of one child per twin, trying to get unique mother info
```{r}
one_twin_only_df = distinct(twins_only, number,.keep_all = TRUE)
nrow(one_twin_only_df) # 53563


# Do not look for discrepancies
for(t in 1:length(twins_only$number)) {
  # While iterating through each twin, if there are discrepancies between mothers data, random choose which row of data to use
  
  # If no discrepancy, add first one to mothers_only df
}

# GOAL: mothers only should have data for all unique mothers in the dataset, easy for us to create bar charts giving an overview of the mothers health
```

RQ1: Difference in birth weight of twin pairs? - Create a dataframe for each twin pair showing their mother's info + difference in birth weight
Visualize mothers data for twins
```{r}
# mager41: mother's age - values 13 to 50 with 13-49 indicating 13-49 years old, 50 indicating 50-64 years old
# meduc: mother's education coded in 9 levels - 1: 8th grade or less; 2: 9th through 12th grade with no diploma; 3: High school graduate or GED completed; 4: Some college credit, but not a degree; 5: Associate degree (AA,AS); 6: Bachelor’s degree (BA, AB, BS); 7: Master’s degree (MA, MS, MEng, MEd, MSW, MBA); 8: Doctorate (PhD, EdD) or Professional Degree (MD, DDS, DVM, LLB, JD); 9: Unknown
# mracerec: mother's race coded in 4 levels: 1: White; 2: Black; 3: American Indian / Alaskan Native; 4: Asian / Pacific Islander
# wtgain: weight gain in pounds - 00-97: Weight gain in pounds; 98: 98 pounds and over; 99: Unknown or not stated
# cig_1: number of cigarettes daily during 1st Trimester - 00-97: Number of cigarettes daily; 98: 98 or more cigarettes daily; 99: Unknown or not stated
# cig_2: number of cigarettes daily during 2nd Trimester - 00-97: Number of cigarettes daily; 98: 98 or more cigarettes daily; 99: Unknown or not stated
# cig_3: number of cigarettes daily during 3rd Trimester - 00-97: Number of cigarettes daily; 98: 98 or more cigarettes daily; 99: Unknown or not stated
# rf_diab: prepregnancy diabetes - Y: Yes; N: No; U: Unknown or not stated
# rf_gest: gestational diabetes - Y: Yes; N: No; U: Unknown or not stated
# rf_phyp: prepregnancy hypertension - Y: Yes; N: No; U: Unknown or not stated
# rf_ghyp: gestational hypertension - Y: Yes; N: No; U: Unknown or not stated
# rf_eclam: hypertension eclampsia - Y: Yes; N: No; U: Unknown or not stated
# rf_ppterm: previous preterm birth - Y: Yes; N: No; U: Unknown or not stated
# rf_ppoutc: poor pregnancy outcome - Y: Yes; N: No; U: Unknown or not stated
# rf_cesar: previous cesareans - Y: Yes; N: No; U: Unknown or not stated
# rf_ncesar: number of previous cesareans - 00: None; 01-30: Number of previous cesareans; 99: Unknown or not stated


# List all the colnames/variables for mothers that we are interested in
mother_attrs = c("mager41", "meduc", "mracerec", "wtgain", "cig_1", "cig_2", "cig_3", "rf_diab", "rf_diab", "rf_gest", "rf_phyp", "rf_ghyp", "rf_eclam", "rf_ppterm", "rf_ppoutc", "rf_cesar", "rf_ncesar")


plot_list = list()
plot_list[[1]] <- ggplot(one_twin_only_df, aes(mager41)) +
                  geom_histogram(binwidth = 1)

plot_list[[2]] <- ggplot(one_twin_only_df, aes(meduc)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[3]] <- ggplot(one_twin_only_df, aes(mracerec)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[4]] <- ggplot(one_twin_only_df, aes(wtgain)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[5]] <- ggplot(one_twin_only_df, aes(cig_1)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[6]] <- ggplot(one_twin_only_df, aes(cig_2)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[7]] <- ggplot(one_twin_only_df, aes(cig_3)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[8]] <- ggplot(one_twin_only_df, aes(rf_diab)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[9]] <- ggplot(one_twin_only_df, aes(rf_diab)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[10]] <- ggplot(one_twin_only_df, aes(rf_gest)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[11]] <- ggplot(one_twin_only_df, aes(rf_phyp)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[12]] <- ggplot(one_twin_only_df, aes(rf_ghyp)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[13]] <- ggplot(one_twin_only_df, aes(rf_eclam)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[14]] <- ggplot(one_twin_only_df, aes(rf_ppterm)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[15]] <- ggplot(one_twin_only_df, aes(rf_ppoutc)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[16]] <- ggplot(one_twin_only_df, aes(rf_cesar)) +
                  geom_histogram(binwidth = 1, stat="count")

plot_list[[17]] <- ggplot(one_twin_only_df, aes(rf_ncesar)) +
                  geom_histogram(binwidth = 1, stat="count")

# BUG: Var is wrong type
# idx = 1
# for (i in 1:length(colnames(one_twin_only_df))) {
#   col = colnames(one_twin_only_df)[i]
#   var = as.name(col)
#   
#   if (col %in% mother_attrs) {
#     p = ggplot(one_twin_only_df, aes(var)) +
#                         geom_histogram(binwidth = 1, stat="count")
#     plot_list[[idx]] <- p
#     idx = idx + 1 # increment
#     
#   }
#   
# }
  
# for (i in 1:length(mother_attrs)) {
#   var = mother_attrs[i]
#   print(var)
#   var = as.name(var)
#   print(var)
#   p <- ggplot(one_twin_only_df, aes(var)) +
#   geom_histogram(binwidth = 1, stat="count")
#   plot_list[[i]] <- p
#   Sys.sleep(2)
# }

# Save plots in one PDF
pdf("mothers_eda.pdf")
for (i in 1:length(mother_attrs)) {
    print(plot_list[[i]])
}
dev.off()

```

What is the sex makeup of twins? 
```{r}
# Process data for pair twins
# Get unique twin pair numbers
unique_pairs = unique(twins_only['number'])
number_twin_sex_df = data.frame(number=integer(0), pair_sex=character(0))
assertthat::are_equal(nrow(unique_pairs), nrow(twins_only)/2.)

for (pair in unique_pairs) {
  twins = twins_only[twins_only$number %in% pair,]
  assertthat::are_equal(nrow(twins), 2)
  
  num_m = 0
  num_f = 0 
  unknown = 0
  
  # Count the number of Males/Females
  twins_sex = twins["sex"]
  print(twins_sex[[1]])
  for (s in twins_sex[[1]]){
    if ("F" %in% s) {
      num_f = num_f + 1
    }
    else if ("M" %in% s) {
      num_m = num_m + 1
    }
    else {
      unknown = unknown + 1
    }
  }

  print(num_f)
  print(num_m)
  # Log the number of Males/Females
  pair_sex = "unknown"
  if (num_f == 2) {
    pair_sex = "2F"
  }
  else if (num_m == 2) {
    pair_sex = "2M"
  }
  else if (num_f == 1 & num_m == 1) {
    pair_sex = "MF"
  }

  tmp<-data.frame(pair, pair_sex)
  names(tmp)<-c("number","pair_sex")
  number_twin_sex_df <- rbind(number_twin_sex_df, tmp)
}

assertthat::are_equal(nrow(number_twin_sex_df), nrow(unique_pairs))

# Visualize the result 
plot_list = list() 
ggplot(number_twin_sex_df, aes("pair_sex")) +
                  geom_histogram(binwidth = 1, stat="count")
# plot_list[[1]] <- p

# Save plots in one PDF
# pdf("twins_sex.pdf")
# for (i in 1:1) {
#     print(plot_list[[i]])
# }
# dev.off()


```

RQ2: Mortality rate? - Create  a df for each twin pair showing their mother's info + categorical encoding for mortality (looking at Death column)

RQ4: Create a df for each twin pair showing mother's info + categorical encoding for chromosomal abnormalities (both, one) for each chromosomal abnormality


```{r}
# Create a new wide dataframe
twins_only_numbers <- data.frame(number=character(),
                                 n_children=integer(),
                                 infant_id_og = character(), 
                                 birth_order = integer(), # originally called "lbo"
                                 infant_num = character(), # child number we create
                                 stringsAsFactors=FALSE) 

# For each unique Number
# If there are 2 children and lbo is 1 and 2
# Get unique twin/child group numbers
children_groups = unique(matched_children_df['number'])

for(i in 1:length(twins_only$number)) {
  curr_number = children_groups$number[i]
  
  children = filter(twins_only,  number == curr_number)
  
  if (length(children$n_children) == 2) {
    twins_only <- bind_rows(twins_only, children, .id=NULL)
  }
  
  if (length(children$n_children) > 2) {
    other <- bind_rows(other, children, .id=NULL)
  }
    
}

head(twins_only)
# Add each value 
```



Count and visualize number of twins vs. triplets vs. ...
```{r}

```

Visualize specific data columns of interest
Most twins in the dataset live (death == 0)
```{r}
ggplot(twins2013, aes(death)) +
  geom_histogram()
```

```{r}
# WIP. Need to figure out how to mutate plot_list in plot_histograms function
# Make plots
plot_histograms <- function(var, plot_list) {
  out <- tryCatch(
    {
      p = ggplot(one_twin_only_df, aes(as.integer(var))) +
          geom_histogram(binwidth = 1)
    },
    error=function(cond) {
      p = ggplot(one_twin_only_df, aes(as.factor(var))) +
          geom_histogram(binwidth = 1, stat="count")
    },
    finally={
      plot_list[[i]] <- p
      print(plot_list)
      sprintf('Done plotting %s', var)
    }
  )
  return(out)
}

plot_list <- lapply(mother_attrs, plot_histograms, plot_list)
typeof(y)
plot_list
```

```{r}
unique_pairs = unique(twins_only['number'])
number_twin_sex_df = data.frame(number=integer(0), pair_sex=character(0))
assertthat::are_equal(nrow(unique_pairs), nrow(twins_only)/2.)

for (i in 1:nrow(unique_pairs)) {
  pair = unique_pairs[i,]
  twins = twins_only[twins_only$number %in% pair,]
  assertthat::are_equal(nrow(twins), 2)
  
  num_m = 0
  num_f = 0 
  unknown = 0
  
  # Count the number of Males/Females
  twins_sex = twins["sex"]
  print(twins_sex[[1]])
  for (s in twins_sex[[1]]){
    if (s == "F"){
      num_f = num_f + 1
    }
    else if (s == "M"){
      num_m = num_m + 1
    }
    
    print(num_f)
  }
}
```