---
title: "match_twins"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

Match the observations to twins according to mother's demographics and birth date 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
# Load data
twins2013<-read.csv("./data/twins2013.csv")
```

```{r}
#procedure proposal
# Question: What does "procedure proposal" mean? -Lol you could ignore this. Was a note for myself since I wrote the procedure first and then wrote the code the next day.

colnames(twins2013)
#combine variables dob_mm, dob_wk, mager41, uprevis, wtgain, cig_1, cig_2, cig_3
  #Note that this could be one limitation in our study since we can't say for sure 100% that we are absolutely right about 
  #the matching. There is still a small prob that there are some moms will have the same values for all those variables.

#mother’s age, state of birth, state of residence, race, and education, marital status of the parents, father’s age, father’s education, and the month in which prenatal care began
twins2013_1<-twins2013%>%mutate(identifier=paste(mager41, restatus, meduc, mracerec, mar, fagerec11, precare, uprevis, cig_1, cig_2, cig_3, dob_mm, dob_wk, fracerec, wtgain, sep=", "))
table(is.na(twins2013_1$identifier))

#see how many unique combinations are there and see if it is approx. half of the obs
# Question: Do we assume that matching on the above identifiers would lead to approx half of the obs because matching would lead to twins (2 entries/rows)? -no we kept the original structure that one row is one kid. The pair could be identified by `number`. 
length(unique(twins2013_1$identifier)) 

#give the same pairs a unique twin number just to make it computationally easier for later steps
pair_num<-as.data.frame(unique(twins2013_1$identifier))
pair_num$number<-row_number(pair_num)
colnames(pair_num)<-c("identifier", "number")
twins_2013_2<-twins2013_1%>%left_join(pair_num, by="identifier")
table(is.na(twins_2013_2$identifier))

#see how many twins, triplets, quadruplets in our data
#2:132589   3:4392      4:270    5:66
table(twins2013$dplural)
numintwin<-twins_2013_2%>%group_by(number)%>%summarize(n_children=n())
check<-twins_2013_2%>%filter(number%in%numintwin[numintwin$n_children==6 ,1]$number)
twins_2013_3<-twins_2013_2%>%left_join(numintwin, by="number")
table(twins_2013_3$n_children)
twins_2013_nosing<-twins_2013_3%>%filter(n_children!=1) # no singletons
write.csv(twins_2013_nosing, "./data/twins2013_matched_children.csv")





DISCARD CODE BELOW - see birth order.Rmd for birth order analysis
-------------------------------------------------------------------------------------------------------------------------------------------------------------------
#unmatched percent
# Question: Where does each number come from? This should be from the table above in line 49
12992/123723
# Question: Where does each number come from? This is just 1 minus the fraction above. I was too lazy doing the calculation myself lol
110731/123723

#give birth order and remove those who can't determine
twins_2013_twins<-twins_2013_3%>%filter(n_children==2)
twins_2013_triplets<-twins_2013_3%>%filter(n_children==3)
twins_2013_quad<-twins_2013_3%>%filter(n_children==4)
twins_2013_quin<-twins_2013_3%>%filter(n_children==5)

twins_2013_quin%>%group_by(number)%>%summarise(orderable=length(unique(lbo))==5)
twins_2013_quin_ord<-twins_2013_quin

orderable_quad<-twins_2013_quad%>%group_by(number)%>%summarise(orderable=length(unique(lbo))==4)
table(orderable_quad$orderable)
falsenum<-with(orderable_quad, orderable_quad[orderable==FALSE,]$number)
twins_2013_quad%>%filter(number%in%falsenum)
twins_2013_quad_ord<-twins_2013_quad%>%filter(!(number%in%falsenum))

orderable_trip<-twins_2013_triplets%>%group_by(number)%>%summarise(orderable=length(unique(lbo))==3)
table(orderable_trip$orderable)
falsenum<-with(orderable_trip, orderable_trip[orderable==FALSE,]$number)
twins_2013_triplets%>%filter(number%in%falsenum)
twins_2013_trip_ord<-twins_2013_triplets%>%filter(!(number%in%falsenum))

orderable_twin<-twins_2013_twins%>%group_by(number)%>%summarise(orderable=length(unique(lbo))==2)
table(orderable_twin$orderable)
falsenum<-with(orderable_twin, orderable_twin[orderable==FALSE,]$number)
temp<-twins_2013_twins%>%filter(number%in%falsenum)
twins_2013_twin_ord<-twins_2013_twins%>%filter(!(number%in%falsenum))

#110731 no singletons
nrow(twins_2013_quad_ord)+nrow(twins_2013_quin_ord)+nrow(twins_2013_trip_ord)+nrow(twins_2013_twin_ord)
# Question: Where does each number come from? This should from the calculation above indicate the proportion of pairs that has order information out of matched pais.
87363/110731
# Question: Where does each number come from? Why is the denominator different from the above line of code? I was comparing the proportion of matched pairs of the number of twins, triplets that the dplural variable in the data tells us while the above is of the number of pairs that we matched
87363/123723

```

# What is the below chunk of code? It looks like an older/newer version. Can you integrate with thee above code somehow so we have only one version of the code? Don't want to accidentally edit/update the wrong version! - I believe this is just a old version when I was trying to see whether a different year of data gives us similar matched proportions. The below is 2011.
```{r}
library(tidyverse)
#twins2011<-read.csv("~/Desktop/twins2011.csv")
```

```{r}
#procedure proposal
colnames(twins2011)
#combine variables dob_mm, dob_wk, mager41, uprevis, wtgain, cig_1, cig_2, cig_3
  #Note that this could be one limitation in our study since we can't say for sure 100% that we are absolutely right about 
  #the matching. There is still a small prob that there are some moms will have the same values for all those variables.

#mother’s age, state of birth, state of residence, race, and education, marital status of the parents, father’s age, father’s education, and the month in which prenatal care began
twins2011_1<-twins2011%>%mutate(identifier=paste(mager41, restatus, meduc, mracerec, mar, fagerec11, precare, dplural, uprevis, cig_1, cig_2, cig_3, dob_mm, dob_wk, fracerec, wtgain, sep=", "))
table(is.na(twins2011_1$identifier))

#see how many unique combinations are there and see if it is approx. half of the obs
length(unique(twins2011_1$identifier)) 
46479/123723 #49.1%

#give the same pairs a unique twin number
pair_num1<-as.data.frame(unique(twins2011_1$identifier))
pair_num1$number<-row_number(pair_num1)
colnames(pair_num1)<-c("identifier", "number")
twins_2011_2<-twins2011_1%>%left_join(pair_num1, by="identifier")
table(is.na(twins_2011_2$identifier))

#see how many twins, triplets, quadruplets in our data
#2:132589   3:4392      4:270    5:66
table(twins2011$dplural)
numintwin1<-twins_2011_2%>%group_by(number)%>%summarise(n_children=n())
table(numintwin1$n_children)
check<-twins_2011_2%>%filter(number%in%numintwin1[numintwin1$n_children==1,1]$number)
check$identifier
#use gather/spread to combine the mortality and birth_order and birth weight info





```

