---
title: "pc_algo"
author: "Eunice Jun"
date: "6/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("Rgraphviz")
BiocManager::install("RBGL")

install.packages("ggm", repos="http://cran.fhcrc.org/")
install.packages("graph", repos="http://cran.fhcrc.org/")
install.packages("pcalg", repos="http://cran.fhcrc.org/")
library("graph")
library("pcalg")

#biocLite("Rgraphviz")  ## seems not to run on many systems
library("Rgraphviz")

plotcpdag <- "Rgraphviz" %in% print(.packages(lib.loc = .libPaths()[1])) 
  
## You may need to run these as well.
install.packages(c("abind", "corpcor", "sfsmisc", "robustbase"),repo ="http://cran.fhcrc.org/")

library(abind)
library(corpcor)
library(sfsmisc)
library(robustbase)
```

Run PC algo to get causal structure of data for individual babies as unit of analysis

Clean up data and make continuous variables categorical
```{r}

```

```{r}
## Adapted from HW 3
matched_children_df<-read.csv("./data/twins2013_matched_children.csv")
twins_only <- filter(matched_children_df, n_children == 2)

# Clean data before passing to PC algo
# Make weight gain categorical
twins_only$wtgain_cat<-cut(twins_only$wtgain, seq(0,99,5), labels=c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19))
# Make cigarette categorical
twins_only$cig_1_binary<-ifelse(twins_only$cig_1>0, 1, 0)
twins_only$cig_2_binary<-ifelse(twins_only$cig_2>0, 1, 0)
twins_only$cig_3_binary<-ifelse(twins_only$cig_3>0, 1, 0)
# Add a categorical variable indicating if baby has any one of the congenital abnormalities
twins_only$ca_combined<-paste0(twins_only$ca_anen, twins_only$ca_menin, twins_only$ca_heart, twins_only$ca_hernia, twins_only$ca_ompha, twins_only$ca_gastro, twins_only$ca_limb, twins_only$ca_cleftlp, twins_only$ca_cleft, twins_only$ca_downs, twins_only$ca_chrom, twins_only$ca_hypos)
twins_only$ca_count<-str_count(twins_only$ca_combined, "Y")
twins_only$ca_binary<-ifelse(twins_only$ca_count>0, 1, 0)
# Make mother's health numeric 
twins_only$rf_diab_num<-ifelse(twins_only$rf_diab=="Y", 1, 0)
twins_only$rf_gest_num<-ifelse(twins_only$rf_gest=="Y", 1, 0)
twins_only$rf_phyp_num<-ifelse(twins_only$rf_phyp=="Y", 1, 0)
twins_only$rf_ghyp_num<-ifelse(twins_only$rf_ghyp=="Y", 1, 0)
twins_only$rf_eclam_num<-ifelse(twins_only$rf_eclam=="Y", 1, 0)
twins_only$rf_ppterm_num<-ifelse(twins_only$rf_ppterm=="Y", 1, 0)
twins_only$rf_ppoutc_num<-ifelse(twins_only$rf_ppoutc=="Y", 1, 0)
twins_only$rf_cesar_num<-ifelse(twins_only$rf_cesar=="Y", 1, 0)
# Make sex numeric
twins_only$sex_num<-ifelse(twins_only$sex=="F",1,0)
# Make newborn abnormal conditions numeric
twins_only$ab_vent_num<-ifelse(twins_only$ab_vent=="Y", 1, 0)
twins_only$ab_nicu_num<-ifelse(twins_only$ab_nicu=="Y", 1, 0)
twins_only$ab_surfac_num<-ifelse(twins_only$ab_surfac=="Y", 1, 0)
twins_only$ab_antibio_num<-ifelse(twins_only$ab_antibio=="Y", 1, 0)

keeps <- c("dob_yy", "dob_mm", "dob_wk", "mager9", "meduc", "mracerec", "fracerec", "fagerec11", "wtgain_cat","cig_1_binary", "cig_2_binary", "cig_3_binary", "mar", "rf_diab_num", "rf_gest_num", "rf_phyp_num", "rf_ghyp_num", "rf_eclam_num", "rf_ppterm_num", "rf_ppoutc_num", "rf_cesar_num", "md_route", "dmeth_rec", "apgar5", "sex_num", "ab_vent_num", "ab_nicu_num", "ab_surfac_num", "ab_antibio_num", "ca_binary", "number", "death")

data<-twins_only[ , names(twins_only) %in% keeps]
assertthat::are_equal(nrow(data), nrow(twins_only))
```


Run the PC algo
```{r}
# Initialize PC algo
p <- length(keeps) # 62 variables 
n <- 107126 # sample size 
rnames <- rownames(data)
names <- rnames <- colnames(data)
suffStat <- list(C=data, n=n)

attr(data,"names")

##### Using the PC Algorithm to do the job for us:
indepTest <- pcalg::gaussCItest

## estimate CPDAG
alpha <- 0.05
pc.fit <- pcalg::pc(suffStat, indepTest, p = p, alpha = alpha, verbose = TRUE)
pcalg::showAmat(pc.fit)
pcalg::showEdgeList(pc.fit,names)
if (plotcpdag) {
  plot(pc.fit, main = "Estimated CPDAG",labels=colnames(data))
}
```

Make the unit of analysis a twin pair
```{r}
matched_children_df<-read.csv("./data/twins2013_matched_children.csv")
twins_only <- filter(matched_children_df, n_children == 2)

# Make mother's health numeric 
twins_only$rf_diab_num<-ifelse(twins_only$rf_diab=="Y", 1, 0)
twins_only$rf_gest_num<-ifelse(twins_only$rf_gest=="Y", 1, 0)
twins_only$rf_phyp_num<-ifelse(twins_only$rf_phyp=="Y", 1, 0)
twins_only$rf_ghyp_num<-ifelse(twins_only$rf_ghyp=="Y", 1, 0)
twins_only$rf_eclam_num<-ifelse(twins_only$rf_eclam=="Y", 1, 0)
twins_only$rf_ppterm_num<-ifelse(twins_only$rf_ppterm=="Y", 1, 0)
twins_only$rf_ppoutc_num<-ifelse(twins_only$rf_ppoutc=="Y", 1, 0)
twins_only$rf_cesar_num<-ifelse(twins_only$rf_cesar=="Y", 1, 0)
# Make weight gain categorical
twins_only$wtgain_cat<-cut(twins_only$wtgain, seq(0,99,5), labels=c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19))
# Make cigarette categorical
twins_only$cig_1_binary<-ifelse(twins_only$cig_1>0, 1, 0)
twins_only$cig_2_binary<-ifelse(twins_only$cig_2>0, 1, 0)
twins_only$cig_3_binary<-ifelse(twins_only$cig_3>0, 1, 0)

# Twin pair sex
twins_sex<-twins_only%>%group_by(number)%>%summarise(sex_pair=paste0(sex[1],sex[2]),
sex_pairb=ifelse(sex_pair=="FF", 0, ifelse(sex_pair %in% c("MF","FM"),1,2)),)

# Twin pair congenital abnormalities
# Add a categorical variable indicating if baby has any one of the congenital abnormalities
twins_only$ca_combined<-paste0(twins_only$ca_anen, twins_only$ca_menin, twins_only$ca_heart, twins_only$ca_hernia, twins_only$ca_ompha, twins_only$ca_gastro, twins_only$ca_limb, twins_only$ca_cleftlp, twins_only$ca_cleft, twins_only$ca_downs, twins_only$ca_chrom, twins_only$ca_hypos)
twins_only$ca_count<-str_count(twins_only$ca_combined, "Y")
twins_only$ca_binary<-ifelse(twins_only$ca_count>0, 1, 0)
twins_ca<-twins_only%>%group_by(number)%>%summarise(ca_pair=paste0(ca_binary[1],ca_binary[2]),
ca_pairb=ifelse(ca_pair=="11", 2, ifelse(ca_pair %in% c("01","10"),1,0)),)

# Twin pair newborn abnormalities
# Make newborn abnormal conditions numeric
twins_only$ab_vent_num<-ifelse(twins_only$ab_vent=="Y", 1, 0)
twins_only$ab_nicu_num<-ifelse(twins_only$ab_nicu=="Y", 1, 0)
twins_only$ab_surfac_num<-ifelse(twins_only$ab_surfac=="Y", 1, 0)
twins_only$ab_antibio_num<-ifelse(twins_only$ab_antibio=="Y", 1, 0)
twins_only$ab_combined<-paste0(twins_only$ab_vent, twins_only$ab_nicu, twins_only$ab_surfac, twins_only$ab_antibio)
twins_only$ab_count<-str_count(twins_only$ab_combined, "Y")
twins_only$ab_binary<-ifelse(twins_only$ab_count>0, 1, 0)
twins_ab<-twins_only%>%group_by(number)%>%summarise(ab_pair=paste0(ab_binary[1],ab_binary[2]),
ab_pairb=ifelse(ab_pair=="11", 2, ifelse(ab_pair %in% c("01","10"),1,0)),)

# Twin pair weight difference
wtdiff<-twins_only%>%group_by(number)%>%summarise(wt_diff=max(dbwt)-min(dbwt))

# Twin pair death
twins_death<-twins_only%>%group_by(number)%>%summarise(death_pair=paste0(death[1],death[2]),
death_pairb=ifelse(death_pair=="11", 2, ifelse(death_pair %in% c("01","10"),1,0)),)


# Merge all the tables together
tmp <- merge(x = twins_sex, y = twins_ca, by = "number")
tmp <- merge(x = tmp, y = twins_ab, by = "number")
tmp <- merge(x = tmp, y = wtdiff, by = "number")
tmp <- merge(x = tmp, y = twins_death, by = "number")
tmp <- merge(x = tmp, y = twins_only, by = "number")

#tmp <- left_join(x = twins_sex, y = twins_ca, by = "number")
#tmp <- left_join(x = tmp, y = twins_ab, by = "number")
#tmp <- left_join(x = tmp, y = wtdiff, by = "number")
#tmp <- left_join(x = tmp, y = twins_death, by = "number")
#tmp <- left_join(x = tmp, y = twins_only, by = "number")

tmp<-distinct(tmp, number, .keep_all = TRUE)
assertthat::are_equal(nrow(tmp)*2, nrow(twins_only))


keeps <- c("dob_yy", "dob_mm", "dob_wk", "mager9", "meduc", "mracerec", "fracerec", "fagerec11", "wtgain_cat","cig_1_binary", "cig_2_binary", "cig_3_binary", "mar", "rf_diab_num", "rf_gest_num", "rf_phyp_num", "rf_ghyp_num", "rf_eclam_num", "rf_ppterm_num", "rf_ppoutc_num", "rf_cesar_num", "md_route", "dmeth_rec", "apgar5", "sex_pairb", "ab_pairb", "ca_pairb", "death_pairb", "wt_diff")

data<-tmp[ , names(tmp) %in% keeps]
assertthat::are_equal(nrow(data)*2, nrow(twins_only))
```

Run the PC algo
```{r}
# Initialize PC algo
p <- length(keeps) # 62 variables 
assertthat::are_equal(53563, nrow(data))
n <- nrow(data) # sample size 
rnames <- rownames(data)
names <- rnames <- colnames(data)
suffStat <- list(C=data, n=n)

attr(data,"names")

##### Using the PC Algorithm to do the job for us:
indepTest <- pcalg::gaussCItest

## estimate CPDAG
alpha <- 0.05
pc.fit <- pcalg::pc(suffStat, indepTest, p = p, alpha = alpha, verbose = TRUE)
pcalg::showAmat(pc.fit)
pcalg::showEdgeList(pc.fit,names)
if (plotcpdag) {
  plot(pc.fit, main = "Estimated CPDAG",labels=colnames(data))
}
```

Try PC algo with unit of analysis as twin pair again with lbw (instead of weight difference), fewer categories for weight gain, mager41 instead of mager9
```{r}
#mother’s race, age, marital status, weight gain, father’s race, twin pair’s sex combinationn assigned at birth.
matched_children_df<-read.csv("./data/twins2013_matched_children.csv")
twins_only <- filter(matched_children_df, n_children == 2)

# Make mother's health numeric and categorical
twins_only$rf_diab_num<-ifelse(twins_only$rf_diab=="Y", 1, 0)
twins_only$rf_gest_num<-ifelse(twins_only$rf_gest=="Y", 1, 0)
twins_only$rf_phyp_num<-ifelse(twins_only$rf_phyp=="Y", 1, 0)
twins_only$rf_ghyp_num<-ifelse(twins_only$rf_ghyp=="Y", 1, 0)
twins_only$rf_eclam_num<-ifelse(twins_only$rf_eclam=="Y", 1, 0)
twins_only$rf_ppterm_num<-ifelse(twins_only$rf_ppterm=="Y", 1, 0)
twins_only$rf_ppoutc_num<-ifelse(twins_only$rf_ppoutc=="Y", 1, 0)
twins_only$rf_cesar_num<-ifelse(twins_only$rf_cesar=="Y", 1, 0)
# Make weight gain categorical
# Filter out missing weight gain data
twins_only<-filter(twins_only, wtgain < 99)
twins_only$wtgain_cat<-cut(twins_only$wtgain, seq(-1,100,25), labels=c(1,2,3,4))
# Make cigarette categorical
twins_only$cig_1_binary<-ifelse(twins_only$cig_1>0, 1, 0)
twins_only$cig_2_binary<-ifelse(twins_only$cig_2>0, 1, 0)
twins_only$cig_3_binary<-ifelse(twins_only$cig_3>0, 1, 0)

# Twin pair sex
twins_sex<-twins_only%>%group_by(number)%>%summarise(sex_pair=paste0(sex[1],sex[2]),
sex_pairb=ifelse(sex_pair=="FF", 0, ifelse(sex_pair %in% c("MF","FM"),1,2)),)

# Twin pair congenital abnormalities
# Add a categorical variable indicating if baby has any one of the congenital abnormalities
twins_only$ca_combined<-paste0(twins_only$ca_anen, twins_only$ca_menin, twins_only$ca_heart, twins_only$ca_hernia, twins_only$ca_ompha, twins_only$ca_gastro, twins_only$ca_limb, twins_only$ca_cleftlp, twins_only$ca_cleft, twins_only$ca_downs, twins_only$ca_chrom, twins_only$ca_hypos)
twins_only$ca_count<-str_count(twins_only$ca_combined, "Y")
twins_only$ca_binary<-ifelse(twins_only$ca_count>0, 1, 0)
twins_ca<-twins_only%>%group_by(number)%>%summarise(ca_pair=paste0(ca_binary[1],ca_binary[2]),
ca_pairb=ifelse(ca_pair=="11", 2, ifelse(ca_pair %in% c("01","10"),1,0)),)

# Twin pair newborn abnormalities
# Make newborn abnormal conditions numeric
twins_only$ab_vent_num<-ifelse(twins_only$ab_vent=="Y", 1, 0)
twins_only$ab_nicu_num<-ifelse(twins_only$ab_nicu=="Y", 1, 0)
twins_only$ab_surfac_num<-ifelse(twins_only$ab_surfac=="Y", 1, 0)
twins_only$ab_antibio_num<-ifelse(twins_only$ab_antibio=="Y", 1, 0)
twins_only$ab_combined<-paste0(twins_only$ab_vent, twins_only$ab_nicu, twins_only$ab_surfac, twins_only$ab_antibio)
twins_only$ab_count<-str_count(twins_only$ab_combined, "Y")
twins_only$ab_binary<-ifelse(twins_only$ab_count>0, 1, 0)
twins_ab<-twins_only%>%group_by(number)%>%summarise(ab_pair=paste0(ab_binary[1],ab_binary[2]),
ab_pairb=ifelse(ab_pair=="11", 2, ifelse(ab_pair %in% c("01","10"),1,0)),)

# Twin pair weight difference
mean_weight<-twins_only%>%group_by(number)%>%summarise(mean_weight=mean(dbwt[1], dbwt[2]))

# Twin pair low birth weight
# Filter out missing data
twins_only<-filter(twins_only, dbwt < 9999)
twins_only$lbwt<- ifelse(twins_only$dbwt < 2500, 1, 0)
twins_lbw<-twins_only%>%group_by(number)%>%summarise(lbw_pair=paste0(x=lbwt[1],lbwt[2]), lbw_pairb=str_count(lbw_pair,"1"))

# Twin pair death
twins_death<-twins_only%>%group_by(number)%>%summarise(death_pair=paste0(death[1],death[2]),
death_pairb=ifelse(death_pair=="11", 2, ifelse(death_pair %in% c("01","10"),1,0)),)


# Merge all the tables together
tmp <- merge(x = twins_sex, y = twins_ca, by = "number")
tmp <- merge(x = tmp, y = twins_ab, by = "number")
tmp <- merge(x = tmp, y = twins_lbw, by = "number")
tmp <- merge(x = tmp, y = twins_death, by = "number")
tmp <- merge(x = tmp, y = twins_only, by = "number")

tmp<-distinct(tmp, number, .keep_all = TRUE)
assertthat::are_equal(nrow(tmp)*2, nrow(twins_only))


keeps <- c("mager41", "meduc", "mracerec", "fracerec", "wtgain_cat","cig_1_binary", "cig_2_binary", "cig_3_binary", "mar", "sex_pairb", "ab_pairb", "ca_pairb", "death_pairb", "lbw_pairb")

data<-tmp[ , names(tmp) %in% keeps]
assertthat::are_equal(nrow(data)*2, nrow(twins_only))

```
Run the PC algo
```{r}
# Initialize PC algo
p <- length(keeps) # 62 variables 
assertthat::are_equal(50372, nrow(data))
n <- nrow(data) # sample size 
rnames <- rownames(data)
names <- rnames <- colnames(data)
suffStat <- list(C=data, n=n)

attr(data,"names")

##### Using the PC Algorithm to do the job for us:
indepTest <- pcalg::gaussCItest

## estimate CPDAG
alpha <- 0.05
pc.fit <- pcalg::pc(suffStat, indepTest, p = p, alpha = alpha, verbose = TRUE)
pcalg::showAmat(pc.fit)
pcalg::showEdgeList(pc.fit,names)
if (plotcpdag) {
  plot(pc.fit, main = "Estimated CPDAG",labels=colnames(data))
}
```